---
title: "Experiment 1"
output: html_notebook
---

# General imports

```{r}
library(jsonlite)
library(tidyverse)
library(ggthemes)
library(lme4)
library(lmerTest)
library(tidyboot)
```

# Experiment 1

## Setup data

### Import turk survey data 

```{r}
exp1_subjInfo <- read_csv('../data/experiment1/planned_sample/turk/subjInfo.csv')
```

### Import message data and implement preregistered exclusions

```{r}
d_annotated <- read_csv('../data/experiment1/planned_sample/chatMessage/messages_with_annotations.csv') 

# nonNativeSpeakerIDs <- unique((tangramSubjInfo %>% filter(nativeEnglish != "yes"))$gameid)
incompleteIDs <- unique((d_annotated %>% group_by(gameid) %>% 
                           filter(length(unique(trialNum)) != 24))$gameid)
confused <- unique((exp1_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
nonNative <- unique((exp1_subjInfo %>% filter(nativeEnglish != 'yes'))$gameid)

# Some speakers violated instructions by relying fully on location in grid 
location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', 
                      '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', 
                      '5684-f09de856-4d64-423c-b342-b30e7325eb0e', 
                      '7600-346d56b5-bd76-406f-befb-6b47e07d5916')
# Some speakers violated instructions by playing this weird taboo game where they gave riddles 
tabooers <- c('2929-d218f724-b45e-416b-af44-5cab5658bad9', #(opposite of white... primary color...)
              '7600-346d56b5-bd76-406f-befb-6b47e07d5916', # MERRY GO ROUND
              '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
badGames <- c(incompleteIDs, nonNative, location_abusers, confused, tabooers)

d <- d_annotated %>%
  mutate(numFeatures = colorMention + shapeMention + textureMention) %>%
  filter(!(gameid %in% badGames)) %>%
  mutate(occlusions = as.factor(occlusions),
        context = as.factor(context))
```

How many recruited, how many excluded, for what reasons, etc?

```{r}
paste0('total games recruited: ', length(unique(d_annotated$gameid)))
paste0('incomplete ', length(incompleteIDs))
paste0('confused ', length(setdiff(confused,incompleteIDs)))
paste0('non-native english ', length(setdiff(nonNative, union(confused, incompleteIDs))))
paste0('violated instructions ', length(setdiff(c(location_abusers, tabooers), union(nonNative, union(confused, incompleteIDs)))))
paste0('final sample ', length(unique(d$gameid)))
```

### Import click data

```{r}
exp1_clicks <- read_tsv('../data/experiment1/planned_sample/clickedObj/clickedObj.csv') %>%
  filter(!(gameid %in% badGames))
```


## Speaker results

### Visualize message length collapsing across all subjects (Fig. 4A)

```{r}
dodge <- position_dodge(width=0.9)
cbPalette = c(rgb(0.8, .8, .8), rgb(.8, .3, .3),  rgb(0, .6, .4), rgb(.4, .4, .4))

d %>%
  group_by(context, hidden) %>%
  tidyboot_mean(column = numRawWords) %>%
  ggplot(aes(x = context, y = empirical_stat, fill = hidden)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
    theme_few() +
    theme(aspect.ratio = 1, text = element_text(size=20)) +
    scale_fill_manual(values=cbPalette) +
    ylab('mean # words') 

ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/num_words.pdf', width = 4, height = 3)
```

### Full mixed-model for effects of condition

```{r}
contrasts(d$context) <- contr.treatment(2, base = 2)
contrasts(d$occlusions) <- contr.treatment(2, base = 2)

summary(lmer(numRawWords ~ context * occlusions + (1 + context + occlusions | gameid), data = d ))
```

### Using feature data instead (Fig. 4B)

```{r}
d.booted <- d %>% 
  gather(feature, mentioned, shapeMention, colorMention, textureMention) %>%
  group_by(context, hidden, feature) %>% 
  tidyboot_mean(column = mentioned) 

d.booted %>%
  mutate(feature = factor(feature, levels = c('shapeMention', 'colorMention', 'textureMention'))) %>%
  ggplot(aes(x = context, y = mean, fill = hidden)) + 
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
    facet_wrap(~ feature) + 
    theme_few() +
    theme(aspect.ratio = 1, text = element_text(size=20))  +
    ylab('% utterances with feature') +
    scale_fill_manual(values=cbPalette) 

ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/by-feature-analysis.pdf', 
       height = 3, width = 9)
```

Note: Random interaction doesn't converge for full model (less variance in these numbers...)

```{r}
contrasts(d$context) <- contr.treatment(2, base = 2)
contrasts(d$occlusions) <- contr.treatment(2, base = 2)
summary(glmer(colorMention ~ occlusions * context + (1 + occlusions + context | gameid), 
              family = 'binomial', data = d, control = glmerControl(optimizer = 'bobyqa')))
```

```{r}
summary(glmer(textureMention ~ occlusions * context + (1 + occlusions + context | gameid), 
              family = 'binomial', data = d, control = glmerControl(optimizer = 'bobyqa')))
```

### How common are errors?

```{r}
## Errors by condition
numErrorsByCondition <- exp1_clicks %>%
  group_by(context, occlusions) %>%
  summarize(totalCorrect = sum(correct), totalErrors = length(correct) - totalCorrect) 

chisq.test(t(numErrorsByCondition[,3:4]))

## Error distribution by pair
exp1_clicks %>%
  group_by(gameid) %>%
  summarize(numErrors = 24- sum(correct)) %>%
  group_by(numErrors) %>%
  tally()
```

# Additional checks & exploratory analyses:

### Visualize basic (occlusion-absent) pragmatic effect for each participant

```{r}
basic.d <- d %>%
  filter(hidden == 'no') %>%
  group_by(gameid, context) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(context, m) %>%
  mutate(diff = close- far,
         ratio = log(close / far))
 
ggplot(basic.d, aes(x = far, y = close)) +
    geom_point() +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,12) +
    xlim(1,12) +
    theme(aspect.ratio = 1, text = element_text(size=20))
```

## Visualize individual differences in num words

Some people used exactly the same number of words on every trial...

```{r}
hidden.d <- d %>%
  group_by(gameid, context, hidden) %>%
  summarize(m = mean(numFeatures)) %>%
  spread(hidden, m) %>%
  mutate(diff = yes- no,
         ratio = log(yes/no)) 

ggplot(hidden.d, aes(x = no, y = yes)) +
    geom_density_2d() +
    facet_wrap(~ context) +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(0,3) +
    ylab('# words (with occlusions)') +
    xlab('# words (without occlusions)') +
    #ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
    xlim(0,3) +
    theme(aspect.ratio = 1, text = element_text(size=20)) 

```

## Visualize histogram of number of features mentioned in each condition

```{r}
d %>%
  mutate(hidden = fct_recode(hidden, occlusion_present = 'yes', occlusion_absent = 'no')) %>%
  ggplot(aes(x = numFeatures)) +
    geom_histogram(binwidth = 1) +
    facet_grid(context ~ hidden) +
    theme_few()
```

## Look at different combinations of color & texture mentions

```{r}
d %>% group_by(context, occlusions) %>% 
  mutate(numTrials = length(text)) %>% 
  group_by(context, occlusions, colorMention, textureMention) %>% 
  summarize(n = length(text)/mean(numTrials)) %>%
  ggplot(aes(x = colorMention, y = textureMention, fill = n)) +
    geom_bin2d(stat = 'identity') +
    stat_bin2d(geom = "text", aes(label = round(n, 2))) +
    scale_fill_gradient(low = "white", high = "red") +
    facet_grid(occlusions ~ context) +
    theme_few()
```

