---
title: "Model-based analyses"
output: html_notebook
---

In this notebook, we make plots and analyze the outputs of our simulations.

# Imports 

```{r}
library(jsonlite)
library(tidyverse)
library(ggthemes)
library(lme4)
library(lmerTest)
library(tidyboot)
```
# Resource-rational simulations (Section 2)

Make Fig. 2

```{r}
costs <- c(0, .05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5)
RR <- read_csv('../modeling/output/fine_RRlistenerOutput.csv') %>%
  mutate(src = 'listener') %>%
  bind_rows(read_csv('../modeling/output/fine_RRspeakerOutput.csv') %>% 
              mutate(src ='speaker')) %>%
  group_by(src, perspectiveCost, alpha, uttCost) %>%
  mutate(maxProb = prob == max(prob)) %>%
  filter(round(perspectiveCost,2) %in% costs)

RR %>%
  ggplot(aes(x = weighting, y = prob, 
             color = perspectiveCost, group = perspectiveCost)) +
  geom_line(alpha = 0.75) +
  geom_point(aes(x = weighting, y = prob, color = perspectiveCost), 
             data = subset(RR %>% filter(maxProb))) +
  facet_grid(~ src) +
  theme_few() +
  xlim(0,1) +
  theme(aspect.ratio = 1) +
  theme(legend.position="top")

ggsave('simulations.pdf')
```
Look at optimal weights for example cost value

```{r}
RR %>%
  filter(perspectiveCost == 0.2) %>%
  filter(maxProb)
```
Look at change points (for Fig.2 caption)

```{r}
RR %>%
  filter(perspectiveCost == 0, src == 'speaker') %>%
  mutate(changePoint = prob != lead(prob)) %>%
  filter(changePoint)
```

# Dynamics simulations (Appendix B)

```{r}
dynamics <- read_csv('../modeling/output/dynamicsOutput.csv')

dynamics %>%
  group_by(numDatapoints) %>%
  filter(prob == max(prob)) %>%
  ggplot(aes(x = numDatapoints, y = weighting)) +
    geom_line() +
    geom_point() +
    theme_few() +
    xlim(0, 50) +
    ylim(0, 1) +
    labs(x = 'length of interaction', y = 'optimal weighting') +
    theme(aspect.ratio = 1) 

ggsave('dynamics.pdf', width = 2, height = 2, units = 'in')
```

# Model comparison (Appendix C)

Define some helper functions

```{r}
library(coda)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
HPDhi<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
HPDlo<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
```

## Pre-process & write out empirical data from Exp 1 in format for bayesian data analysis

Need to add column annotating whether the critical distractor in close trials differs from target in texture, color, or both...

```{r}
exp1_subjInfo <- read_csv('../data/experiment1/planned_sample/turk/subjInfo.csv')
d_annotated <- read_csv('../data/experiment1/planned_sample/chatMessage/messages_with_annotations.csv') 
incompleteIDs <- d_annotated %>% 
  group_by(gameid) %>%  
  filter(length(unique(trialNum)) != 24) %>%
  pull(gameid) %>%
  unique()

confused <- exp1_subjInfo %>% 
  filter(understandsInstructions != 'yes') %>%
  pull(gameid) %>%
  unique()

nonNative <- exp1_subjInfo %>% 
  filter(nativeEnglish != 'yes') %>%
  pull(gameid) %>%
  unique()

# Some speakers violated instructions by relying fully on location in grid 
location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', 
                      '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', 
                      '5684-f09de856-4d64-423c-b342-b30e7325eb0e', 
                      '7600-346d56b5-bd76-406f-befb-6b47e07d5916')

# Some speakers violated instructions by playing this weird taboo game where they gave riddles 
tabooers <- c('2929-d218f724-b45e-416b-af44-5cab5658bad9', #(opposite of white... primary color...)
              '7600-346d56b5-bd76-406f-befb-6b47e07d5916', # MERRY GO ROUND
              '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???

badGames <- c(incompleteIDs, nonNative, location_abusers, confused, tabooers)

d <- d_annotated %>%
  mutate(numFeatures = colorMention + shapeMention + textureMention) %>%
  filter(!(gameid %in% badGames)) %>%
  mutate(occlusions = as.factor(occlusions),
        context = as.factor(context))

exp1_clicks <- read_tsv('../data/experiment1/planned_sample/clickedObj/clickedObj.csv') %>%
  mutate(correct = ifelse(correct == 'true', 1, 0)) %>%
  filter(!(gameid %in% badGames))

exp1_clicks %>% 
  mutate(names = substr(names,2,nchar(names)-1)) %>%
  separate(names, sep = ',', into = c('item1','item2','item3', 'item4', 'item5'),
           extra = "merge", fill = "left") %>%
  select(-item5) %>%
  gather(distractorNum, distractorName, item1:item4) %>%
  separate(intendedName, into = c('targetTexture', 'targetColor', 'targetShape')) %>%
  filter(!is.na(distractorName)) %>%
  separate(distractorName, into = c('distractorTexture', 'distractorColor', 'distractorShape')) %>%
  filter(distractorShape == targetShape) %>%
  mutate(textureMatch = distractorTexture == targetTexture) %>%
  mutate(colorMatch = distractorColor == targetColor) %>%
  mutate(distractorType = ifelse(textureMatch, 'texture_shape', ifelse(colorMatch, 'color_shape', 'shape_only'))) %>%
  select(gameid, trialNum, distractorType) %>%
  right_join(exp1_clicks) %>%
  mutate(distractorType = ifelse(is.na(distractorType), 'diff_shape', distractorType)) %>%
  right_join(d) %>% 
  select(gameid, trialNum, distractorType, context, hidden, 
         numDistractors, colorMention, shapeMention, textureMention) %>%
  write_csv('../modeling/input/bdaInput.csv')
```

Get empirical data in a format to make predictives plots against model output

```{r}
empiricalData <- d %>% 
  group_by(context, hidden) %>% 
  tidyboot_mean(column = numFeatures) %>%
  ungroup() %>%
  mutate(context = as.character(context)) %>%
  mutate(source = as.character('empirical data')) %>% 
  select(-empirical_stat, -n)
```

## Generate Predictives (Fig. C1)

Look at pattern of predictives for each model, compared to empirical number of features mentioned

```{r}
dodge <- position_dodge(width=0.9)

read.csv('../modeling/output/occlusionSensitiveBDAPredictives.csv') %>% 
  mutate(source = 'occlusion-sensitive model') %>%
  rbind(read.csv('../modeling/output/egocentricBDAPredictives.csv') %>% 
        mutate(source = 'occlusion-blind model')) %>%
  rbind(read.csv('../modeling/output/mixtureBDAPredictives.csv') %>% 
        mutate(source = 'mixture model')) %>%
  mutate(i = row_number()) %>%
  gather(condition, value, -prob, -MCMCprob, -source, -i) %>%
  separate(condition, sep = '_', into = c('context', 'numDistractors', 'distractorType', 'hidden')) %>%
  group_by(context, hidden, source) %>%
  summarize(mean = mean(value),
            ci_lower = round(HPDlo(value), 3),
            ci_upper = round(HPDhi(value), 3)) %>%
  ungroup() %>%
  mutate(hidden = ifelse(hidden == 'hidden', 'yes', 'no')) %>%
  rbind(empiricalData) %>%
  mutate(source = fct_relevel(source, 'empirical data', 'mixture', 
                              'occlusion-sensitive model', 'occlusion-blind model')) %>%
  ggplot(aes(x = context, fill = hidden, y = mean)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
    facet_wrap(~ source) +
    theme_few() +
    ylim(0,3) +
    theme(aspect.ratio = .57, text = element_text(size=20)) +
    ylab('mean # features produced') 

ggsave('predictives.pdf', useDingbats=FALSE )
```
## Determine best-fitting model (Table C1)

```{r}
read_csv('../modeling/output/AIS2.csv') %>%
  group_by(model) %>%
  summarize(marginal = mean(value))
```

Check out inferred mixture weight in mixture model

```{r}
read.csv('../modeling/output/mixtureBDAParams.csv') %>%
  gather(parameter, value, alpha:shapeCost) %>%
  group_by(parameter) %>%
  summarize(mode = estimate_mode(value),
            md_lo = HPDlo(value),
            md_hi = HPDhi(value))
```

Get MAP estimates and CIs from param posteriors for best-fitting occlusion-sensitive model with $w_S = 1$. 

```{r}
params <- read.csv('../modeling/output/occlusionSensitiveBDAParams.csv') 

params %>%
  select(-ownWeighting) %>%
  gather(parameter, value, alpha:shapeCost) %>%
  group_by(parameter) %>%
  summarize(mode = estimate_mode(value),
            md_lo = HPDlo(value),
            md_hi = HPDhi(value))
```

Visualize param posteriors (Fig. S2)

```{r}
numSamples = 1000
params.samples <- params[rep(row.names(params), exp(params$prob)*numSamples), ] %>%   
  select(-ownWeighting) %>%
    gather(parameter, value, alpha:shapeCost)

paramPosteriors = ggplot(params.samples, aes(x=log(value)))+
    geom_density(aes(y=..density..),
                 data =subset(params.samples, parameter == "alpha" ),
                 adjust = 3, alpha=.2, fill="#FF6666")+
    geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter == "textureCost"),
                 adjust = 3, alpha=.2, fill="#FF6666")+
    geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter == "colorCost"),
                 adjust = 3, alpha=.2, fill="#FF6666")+
    geom_density(aes(y=..density..),
                 data=subset(params.samples, parameter == "shapeCost"),
                 adjust = 3, alpha=.2, fill="#FF6666")+
    xlab('log parameter value') +
    ggtitle("Parameter Posteriors (1000 samples)") +
    facet_grid(parameter ~ . , scales = 'free') +
    theme_few()

paramPosteriors
ggsave('paramPosterior.pdf',
       height = 4, width = 6)
```

# Bonus (unused) analyses

Look at big grid of RR analysis for other alpha/uttCost params (need to run with 'coarse-grid' first)

```{r}
bind_rows(
  read_csv('../modeling/output/new_RRlistenerOutput.csv') %>% mutate(src = 'listener'),
  read_csv('../modeling/output/new_RRspeakerOutput.csv') %>% mutate(src ='speaker')
) %>%
  filter(perspectiveCost < 0.55) %>%
  filter(alpha == 2.12) %>%
  group_by(src, perspectiveCost, alpha, uttCost) %>%
  ggplot(aes(x = weighting, y = prob, 
             color = perspectiveCost, group = perspectiveCost)) +
  geom_line() +
  facet_grid(uttCost ~ src) +
  theme_few() +
  xlim(0,1) +
  theme(aspect.ratio = 1)
```

Look at scatter plots of predictives.

```{r}
to.plot <- rbind(read.csv('../modeling/bdaOutput/occlusionSensitivityPredictives.csv') %>% mutate(source = 'occlusion-sensitive model'),
     read.csv('../modeling/bdaOutput/egocentricPredictives.csv') %>% mutate(source = 'occlusion-blind model')
     ) %>%
  mutate(i = row_number()) %>%
  gather(condition, value, -prob, -MCMCprob, -source, -i) %>%
  group_by(condition, source) %>%
  summarize(mean = mean(value),
            ci_lower = round(HPDlo(value), 3),
            ci_upper = round(HPDhi(value), 3)) %>%
  separate(condition, sep = '_', into = c('context', 'numDistractors', 'distractorType', 'hidden')) %>%
  mutate(hidden = ifelse(hidden == 'hidden', 'yes', 'no')) %>%
  rbind(read_csv('../modeling/bdaInput/bdaInput.csv') %>% 
          mutate(numFeatures = color + shape + texture) %>%
          group_by(context, distractorType, numDistractors, hidden) %>% 
          tidyboot_mean(column = numFeatures) %>%
          ungroup() %>%
          mutate(numDistractors = numDistractors - 2) %>%
          mutate(context = as.character(context)) %>%
          mutate(distractorType = gsub("_", "", distractorType)) %>%
          mutate(source = as.character('empirical data')) %>% 
          ungroup() %>%
          select(-empirical_stat, -n)) %>%
  mutate(source = factor(source, levels = c('empirical data', 'occlusion-sensitive model', 
                                            'occlusion-blind model'))) %>%
  unite(col = asdf, mean, ci_lower, ci_upper) %>%
  spread(source, asdf) %>%
  separate(`empirical data`, sep = '_', into = c('emp_mean', 'emp_lower', 'emp_upper'))%>%
  separate(`occlusion-sensitive model`, sep = '_', into = c('occ_mean', 'occ_lower', 'occ_upper')) %>%
  separate(`occlusion-blind model`, sep = '_', into = c('ego_mean', 'ego_lower', 'ego_upper')) %>%
  mutate(emp_mean = as.numeric(emp_mean), emp_lower = as.numeric(emp_lower), emp_upper = as.numeric(emp_upper),
         occ_mean = as.numeric(occ_mean), occ_lower = as.numeric(occ_lower), occ_upper = as.numeric(occ_upper),
         ego_mean = as.numeric(ego_mean), ego_lower = as.numeric(ego_lower), ego_upper = as.numeric(ego_upper)) 

ggplot(to.plot, aes(x = occ_mean, y = emp_mean)) +
  geom_point() +#(stat = 'identity', position = dodge) +
  geom_errorbar(aes(ymin = emp_lower, ymax = emp_upper), width = 0) +
  #geom_errorbarh(aes(xmin = occ_lower, xmax = occ_upper), width = 0) +
  geom_smooth(method = 'lm') +
  theme_few() +
  theme(aspect.ratio = 1, text = element_text(size=20)) +
  xlab('predicted # features')+
  ylab('empirical # features') 

ggsave('../../pragmatics_of_perspective_taking_tex/cognition_format/figures/predictiveScatter.pdf',
       useDingbats=FALSE )
```

```{r}
ggplot(to.plot, aes(x = ego_mean, y = emp_mean)) +
  geom_point() +#(stat = 'identity', position = dodge) +
  geom_errorbar(aes(ymin = emp_lower, ymax = emp_upper), width = 0) +
  #geom_errorbarh(aes(xmin = ego_lower, xmax = ego_upper), width = 0) +
  geom_smooth(method = 'lm') +
  theme_few() +
  theme(aspect.ratio = 1, text = element_text(size=20)) +
  xlab('predicted # features')+
  ylab('empirical # features') 
ggsave('../../pragmatics_of_perspective_taking_tex/cognition_format/figures/egocentricScatter.pdf',
       useDingbats=FALSE)
       #height = 5, width = 9)
```